---
title: "DADA2 Installation"
---

-----------------------------

## **Install binaries from Bioconductor**

Binary installation is available through [the Bioconductor package repository](https://bioconductor.org/). 
Binary installation of the current release version (1.36.0) requires R 4.5 (or later) and Bioconductor version 3.21:

```R
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("dada2")
```

Allow installation of any additional packages requested. You are ready to start using dada2!

*If the above installed an early version of dada2 (e.g. 1.4 or 1.6) you can get the current dada2 release by updating to [Bioconductor 3.21](https://bioconductor.org/news/bioc_3_21_release/): `BiocManager::install(version = "3.21")` then `BiocManager::install()`. This may require updating R as well.*

-----------------------------

## **Install from source**

Source installation is available for R 3.2 or later, and the latest and greatest features will be available first through source installs of the development branch.

### 0. Download

* Current Release [(1.36.0)](https://github.com/benjjneb/dada2/releases): [Zipped](https://github.com/benjjneb/dada2/archive/v1.36.0.zip) or [tarball](https://github.com/benjjneb/dada2/archive/v1.36.0.tar.gz).

* Development Branch [(latest development version)](https://github.com/benjjneb/dada2): [Zipped](https://github.com/benjjneb/dada2/archive/master.zip) or [tarball](https://github.com/benjjneb/dada2/archive/master.tar.gz).

### 1. Install External Dependencies

The following instructions assume that you have installed R 
and have installed 
[Xcode](https://developer.apple.com/xcode/downloads/)
(Mac) or 
[Rtools](http://cran.r-project.org/bin/windows/Rtools/)
(Windows); as well as 
[Bioconductor](http://bioconductor.org/install/).

For Bioconductor,
start a fresh R session, and enter the following.

```{r bioc-install, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(suppressUpdates = FALSE)
BiocManager::install("ShortRead", suppressUpdates = FALSE)
```

### 2. Install using devtools

The easiest method is to use the devtools package to install from github:

```{r devtools-install, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("devtools")
library("devtools")
devtools::install_github("benjjneb/dada2")
```

If that worked you're done (see step 4)! If it failed on missing dependencies, goto step 2.

### 2a. Install manually from source

Alternatively, you can install from the source by hand. First [download the zipped package](download.html) and unzip it (or expand the tarball).

Start a fresh R session, and enter the following.

```{r install-packages, eval=FALSE}
install.packages("path/to/dada2",
                 repos = NULL,
                 type = "source",
                 dependencies = c("Depends", "Suggests","Imports"))
```

For example, if the dada2 source code directory was in `~/github/dada2`,
then the following would work.

```{r install-github-example, eval=FALSE}
install.packages("~/github/dada2",
                 repos = NULL,
                 type = "source",
                 dependencies = c("Depends", "Suggests","Imports"))
```

Check Package Version

```{r packageVersion}
packageVersion("dada2")
```

### 3. Troubleshoot Dependencies (Optional)

At this point, 
there may be complaints about missing dependencies.
To install missing dependencies on either
[CRAN](http://cran.r-project.org/)
or
[Bioconductor](http://bioconductor.org/install/), 
start a fresh R session, and enter the following.

```{r bioc-install-missing, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("missing_package_1")
BiocManager::install("missing_package_2")
# ... and so on
```

### 4. Re-attempt dada2 Installation

Install again, after dependencies have been installed.

Devtools method:
```{r devtools-install-rev2, eval=FALSE}
library("devtools")
devtools::install_github("benjjneb/dada2")
```

Manual install from source method:
```{r install-packages-rev2, eval=FALSE}
install.packages("path/to/dada2",
                 repos = NULL,
                 type = "source",
                 dependencies = c("Depends", "Suggests","Imports"))
```

You should now be done, if installation was successful. If not, do not panic, and post a message to [the issue tracker](https://github.com/benjjneb/dada2/issues).

### 5. Load Package, Explore Function Documentation

```{r load-dada2, message=FALSE}
library("dada2")
```

Now that the package is loaded, you can explore the documentation using the R help system.

```{r documentation-example, eval=FALSE, message=FALSE}
help(package="dada2")
?derepFastq
?dada
```

You are ready to use dada2!